### 2016-05-09 Article/calibration.tex
adding different file for article

\section{Calibration and initialization}

A toi de jouer Vivien...