### 2016-05-09 Article/conclusion.tex

\section{Conclusion}

notre library est trop bien...

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
This work was supported by the ARAP program of ASTAR and by the university Sorbonne UPMC.

\bibliographystyle{bmc-mathphys} 
\bibliography{bmc_article}

\end{backmatter}