### 2016-05-09 Article/introduction.tex

\begin{abstractbox}

\begin{abstract} % abstract

The abstract goes here
\end{abstract}

\begin{keyword}
\kwd{IMU}
\kwd{Sensor Fusion}
\kwd{Sensor Calibration}
\end{keyword}

\end{abstractbox}

\end{frontmatter}

\section{Introduction}