### 2016-05-10 Article/main.tex
modif main.tex

\documentclass[twocolumn]{bmcart}

%%%%%%%%%%%%
% Packages
%%%%%%%%%%%%

\usepackage[utf8]{inputenc} %unicode support

\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{listings}
\usepackage{color}

%\def\includegraphic{}
%\def\includegraphics{}

%%%%%%%%%%%%
% Local definition
%%%%%%%%%%%%

\startlocaldefs

%Definition d un environnement remarque auto-indexe

\newcounter{rmq}[section]
\setcounter{rmq}{0}
\newenvironment{remarque}{\addtocounter{rmq}{1}\textbf{Remarque \thermq:}}

% quaternion style
\def\mathbi#1{\textbf{\em #1}}

% roman letter
\makeatletter
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

% colors
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

% code
\lstset{frame=tb,
  language=C++,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\endlocaldefs

%%%%%%%%%%%%
% Begin Document
%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%
% Title
%%%%%%%%%%%%

\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},     
   noteref={n1},                        % id's of article notes, if any
   email={nizar.ouarti@ipal.cnrs.fr}   % email address
]{\inits{NO}\fnm{Nizar} \snm{Ouarti}}
\author[
   addressref={aff1},   corref={aff1},  
   email={braud.thomas@ipal.cnrs.fr}
]{\inits{TB}\fnm{Thomas} \snm{Braud}}
\author[
   addressref={aff1},   corref={aff1},  
   email={vivien.billaud@ipal.cnrs.fr}
]{\inits{VB}\fnm{Vivien} \snm{Billaud}}

\address[id=aff1]{%                           % unique id
  \orgname{Image Pervasive Access Lab (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
  \city{Singapore}
  }

\end{fmbox}% comment this for two column layout

 \title{OpenMotion: An open source library for attitude estimation.}

\maketitle

%%%%%%%%%%%%
% Abstract + Introduction
%%%%%%%%%%%%

\input{introduction.tex}

%%%%%%%%%%%%
% General Presentation of the library
%%%%%%%%%%%%

\input{prez_library.tex}

%%%%%%%%%%%%
% Presentation of the calibration process
%%%%%%%%%%%%

\input{calibration.tex}

%%%%%%%%%%%%
% State of the art
%%%%%%%%%%%%

\input{state_of_the_art.tex}

%%%%%%%%%%%%
% Conclusion - Discussion
%%%%%%%%%%%%

\input{conclusion.tex}

\end{document}